---
title: "Bridges and Labor Assignment"
author: "Jillian Schat"
date: "2/26/2020"
output: html_document
---

```{r}

# retrieving ME unemployment data


library(blscrapeR)
library(tidyverse)
library(dplyr)
library(data.table)
library(magrittr)
library(ggplot2)

# preparing unemployment data

df <- get_bls_county()
ME_labor = df %>% filter(fips_state == 23)
View(ME_labor)

# retrieving WI bridge data

"https://www.fhwa.dot.gov/bridge/nbi/2012/delimited/ME18.txt"
state = "ME"
year = "18"
URL = paste0("https://www.fhwa.dot.gov/bridge/nbi/20",
       year,
       "/delimited/",
       state,
       year,
       ".txt")
ME_bridges = fread(URL)

# now making fips column in bridges data

br = ME_bridges %>%
        as_tibble() %>%
        select(STATE_CODE_001,
               COUNTY_CODE_003,
               contains("COND")) %>%
        mutate(fips = STATE_CODE_001*1000 + COUNTY_CODE_003)

View(br)

# joining labor data with bridges data

ME_labor$fips
# b/c ME fips data is a character
MEfips <- as.numeric(ME_labor$fips)
ME_labor$fips <- MEfips
ME_labor$fips

?left_join

data <- merge(x = ME_labor, y = br, all = TRUE)
View(data)

data_lm <- lm(data$labor_force ~ data$DECK_COND_058 + data$SUPERSTRUCTURE_COND_059 + 
                      data$SUBSTRUCTURE_COND_060)
summary(data_lm)

```

The multiple R-squared value is 0.0511, and the adjusted R-squared value is 0.04102 with a p value of 
p = 1.319e-15. While highly significant, this indcates that the combined variables superstructure
condition, substructure condition, and deck condition in 2018 only explain at best 5.11% of the variation in labor force in Maine in 2019. It is possible that the condition of the bridge was not the deciding factor when labor forces were allocated in Maine. It is possible that other variables, such as daily traffic, gave less deteriorated bridges higher priority. If this is the case, a greater proportion of the labor force would be allocated to less deteriorated bridges and traffic would be the variable that best explains labor force allocation.

```{r}

data_lm_rate <- lm(data$unemployed_rate ~ data$DECK_COND_058 + data$SUPERSTRUCTURE_COND_059 + 
                      data$SUBSTRUCTURE_COND_060)

summary(data_lm_rate)
```

The multiple R-squared value is 0.04206, and the adjusted R-squared value is 0.03188 with a p value of 1.464e-11. While highly significant, at most only 4.206% of the variation in unemployment rate in Maine in 2019 is predicted by deck condition, superstructure condition, and substructure condition in 2018. It is possible other factors are influencing unemployment rate more than these variables. 

```{r}
data_lm_ <- lm(data$unemployed ~ data$DECK_COND_058 + data$SUPERSTRUCTURE_COND_059 + 
                      data$SUBSTRUCTURE_COND_060 + 1480)
summary(data_lm_)

data_lm_rate_ <- lm(data$unemployed_rate ~ data$DECK_COND_058 + data$SUPERSTRUCTURE_COND_059 + 
                      data$SUBSTRUCTURE_COND_060 + 2.7)
summary(data_lm_rate_)
```

